---
title: "Cellbench heatmap of top differential genes"
author: "Hani Kim"
date: "`r paste0('Initiated on 2020 Oct 31, compiled on ', format(Sys.time(), '%Y %b %d'))`"
output:
  html_document:
    code_folding: hide
    fig_height: 12
    fig_width: 12
    toc: yes
    number_sections: true
    toc_depth: 3
    toc_float: yes
    theme: paper
editor_options: 
  chunk_output_type: console
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Summary
This script generates heatmap of gene expression for the top 200 differential genes identified for each cell type by the differential analysis methods. 

# Loading 

## Packages
```{r, message = FALSE}
suppressPackageStartupMessages({
  library(SingleCellExperiment)
  library(pheatmap)
  library(ggplot2)
  library(dplyr)
})
```

## Load 3-class cellbench data and dx results
```{r}
load("../../../Data/cellbench_sce_3class.RData", verbose=T)
rm(sce_CELseq2_log)
rm(sce_Dropseq_log)

load("../../../Data/cellbench_DE_benchmark_results.RData", verbose=T)
load("../../../Data/cellbench_Seurat_benchmark_results.RData", verbose=T)
load("../../../Data/cellbench_additionalMethods_benchmark_results.RData", verbose=T)
```

## Source functions
```{r}
source("../../../R/utils.R")
source("../../../R/doDE.R")
```

# Extract genes
```{r}
DE.res <- list(Limma=Limma.res,
               Voom=Voom.res,
               MAST=MAST.res,
               ttest=ttest.res,
               Wilcoxon=Wilcoxon.res,
               EdgeR=EdgeR.res,
               LR = LR.res,
               ROC = ROC.res
               )

DE.gene <- lapply(DE.res, function(x) lapply(x, function(y) names(y)))
# sapply(1:3, function(i) sapply(DE.gene, function(x) length(x[[i]])) == 13542)
```

# Generate Upset plot of overlaps (top 50 genes)
```{r}
cty <- names(DE.gene[[1]])
lapply(cty, function(celltype) {
    DE.gene.celltype <- lapply(DE.gene, function(x) x[[celltype]][1:50])
    names(DE.gene.celltype) <- names(DE.gene)
    pdf(file = paste0("../../../Figures/cellbench/cellbench_upset_DE_",celltype,".pdf"))
    print(UpSetR::upset(UpSetR::fromList(DE.gene.celltype), nsets=11))
    dev.off()
})
```

Gnerate list for RankCorr and scGeneFit
```{r}
####### rankCorr
rankCorrGenes = rankCorrRes[["140"]]
exprsMat = logcounts(sce_10x_log)
res = doLimma(exprsMat[rownames(exprsMat) %in% rankCorrGenes,],
              sce_10x_log$celltype)
res = getStats(res, method = "Limma")
rankCorrGenes.res = lapply(res, function(x) names(x[1:200]))
table(table(unlist(rankCorrGenes.res)) == 1)

scGeneFitGenes = scGeneFitRes[["600"]]
exprsMat = logcounts(sce_10x_log)
res = doLimma(exprsMat[rownames(exprsMat) %in% scGeneFitGenes,],
              sce_10x_log$celltype)
res = getStats(res, method = "Limma")
scGeneFitGenes.res = lapply(res, function(x) names(x[1:200]))
table(table(unlist(scGeneFitGenes.res)) == 1)
```

# Generate heatmaps (top 200 genes)
```{r}
#### select top 200 genes
DE.top <- lapply(DE.gene, function(x) lapply(x, function(y) y[1:200]))
DE.top$RankCorr = rankCorrGenes.res
DE.top$scGeneFit = scGeneFitGenes.res
  
#### reorder cell types
celltype <- as.character(sce_10x_log$celltype)
celltype[celltype %in% "HCC827"] <- "1_HCC827"
celltype[celltype %in% "H2228"] <- "2_H2228"
celltype[celltype %in% "H1975"] <- "3_H1975"

counter <- 0
lapply(DE.top, function(top) {
    counter <<- counter + 1
    method_name <- names(DE.top)[[counter]]
    top <- top[c("HCC827", "H2228", "H1975")]
    
    plotHeatmap(sce_10x_log, top, celltype) %>% ggsave(filename=paste0("../../../Figures/cellbench/cellbench_heatmap_DE_", method_name, ".pdf"), width=6,height=5)
})

counter <- 0
lapply(DE.top, function(top) {
    counter <<- counter + 1
    method_name <- names(DE.top)[[counter]]
    top <- top[c("HCC827", "H2228", "H1975")]
    
    plotHeatmap(sce_10x_log, top, celltype, slim=T) %>% ggsave(filename=paste0("../../../Figures/cellbench/cellbench_heatmap_DE_", method_name, ".png"), width=6,height=5)
})
```

# Session Info
```{r}
sessioninfo::session_info()
```
